---
title: "Attachment and Healthy Relationship Outcomes"
output: word_document
date: '2022-05-02'
---

```{r setup, include = FALSE}
library(tidyverse)
library(tidytext)
library(textdata)
library(easystats)
knitr::opts_chunk$set(echo = TRUE, include = TRUE, fig.width = 5, fig.height = 4, fig.align = "center")
```


```{r load and rename data}

layperson <- haven::read_spss(
  "Healthy Relationships with Prolific Demo 10.9.19.sav"
)

```

# Layperson's perception of relationship health


This is a project I did for the SPSP 2022 conference investigating the relationship between attachment style and relationship outcomes. I will walk you through the data analysis process.

## Data wrangling 

First, I mutated the data to clarify variable names and make them easier to work with. 


```{r wrangle create new ECR variables}

dat_layperson <- mutate(layperson, ECR1 = ECR_ecr1,
                        ECR2 = ECR_ecr2,
                        ECR3 = ECR_ecr3,
                        ECR4 = ECR_ecr4,
                        ECR5 = ECR_ecr5, 
                        ECR6 = ECR_ecr6, 
                        ECR7 = ECR_ecr7, 
                        ECR8 = ECR_ecr8, 
                        ECR9 = ECR_ecr9
                        )
```

Then, I reverse scores the appropriate items. 

```{r ECR reverse score}

dat_layperson <- dat_layperson |>
  mutate(ECR1 = 8 - ECR1, 
         ECR2 = 8 - ECR2, 
         ECR3 = 8 - ECR3, 
         ECR4 = 8 - ECR4) 

```

I averaged the items across each category to get the total average scores for insecure attachment, avoidant attachment, and anxious attachment styles per participant. 

```{r Sum ECR score}

dat_layperson <- dat_layperson |>
  mutate(InsecureTotalMean = (ECR1 + ECR2 + 
           ECR3 + ECR4 + ECR5 + 
           ECR6 + ECR7 + ECR8 + ECR9)/9) 
    

```

```{r sum ECR for Avoid}

dat_layperson <- dat_layperson |>
  mutate(AvoidTotalMean = (ECR1 + ECR2 + 
           ECR3 + ECR4 + ECR5 + ECR6)/6) 

```

```{r sum ECR for Anxious}

dat_layperson <- dat_layperson |>
  mutate(AnxiousTotalMean = (ECR7 + ECR8 + 
           ECR9)/3)

```

Lastly I created an interaction the anxious and avoidant means. 

```{r Anxious*Avoidant Mean}

dat_layperson <- dat_layperson |> 
  mutate(AvoidxAnxious = AnxiousTotalMean*AvoidTotalMean)

```

```{r save cleaned data output}

write_csv(dat_layperson, "dat_layperson.csv")

```


## Fitting models 

Below, I fit models to check for some of the outcome variables (satisfaction, mutuality, caring, and core similarity) predicted by attachment avoidance, anxiety, and an interaction between the two. I ran the models and checked the parameters for each. 

## Attachment predicting satisfaction

### model 

```{r fit mod satisfaction}
m_satisfaction <- lm(
  satisfaction_1 ~ AnxiousTotalMean + AvoidTotalMean + AnxiousTotalMean*AvoidTotalMean, 
  data = dat_layperson
)
``` 

### parameters

```{r parameters m_sat}

params_satisfaction<- parameters( 
  m_satisfaction, 
  dispersion = TRUE, 
  component = "all", 
  test = NULL, 
  diagnostic = NULL, 
  prior = NULL 
) 
 
print(params_satisfaction) 

```

### Findings

According to the parameters, anxious attachment style is a negative predictor of perception of satisfaction's influence on relationship health. Every 1 unit increase in attachment anxiety corresponds to a -2.35 decrease in perceived healthiness of satisfaction. There was a moderately sized compatibility interval that does not include 0 [-4.44, -.27], meaning there is evidence that this relationship would be seen in the population. Avoidant attachment style is also a negative predictor of perception of satisfaction on relationship healthy. Every 1 unit increase in attachment avoidance corresponds to a -6.47 decrease in perceived healthiness of satisfaction. The compatibility interval [-9.62, -3.32] does not include 0, and we would not be surprised to see this relationship in the population. Interestingly, there was a significant interaction, CI [0.22, 1.84], that trended positively (1.03). Graphing this data will provide more information on the relationship. 

### plots

While I could have done estimate expectations, I am not as familiar with interpreting interactions with two continuous variables. Instead, I split the anxious term into segments of low and high, as below, and saved it into a new data file to use for the plots

```{r splitting anxious into 2}
describe_distribution(dat_layperson$AnxiousTotalMean)


  
dat_layperson1 <- dat_layperson |>
  mutate(
    AnxiousTotalMeanInt = case_when(
      AnxiousTotalMean <= 1.39 ~ "low anxiety",
      AnxiousTotalMean >= 4.61 ~ "high anxiety"
    )
  )


##Lower - 1.39
##Higher - 4.61

```


```{r mod sat plot}
plot_satisfaction <- dat_layperson1 |>
  filter(!is.na(AnxiousTotalMeanInt)) |>
  ggplot() +
  aes(x = AvoidTotalMean, 
      y = satisfaction_1, 
      group = AnxiousTotalMeanInt, 
      color = AnxiousTotalMeanInt) +
  geom_point(shape=21,
             color="darkgrey",
             fill= "darkgrey", 
             size=2, 
             alpha = .5) +
  geom_smooth(method=lm) +
  scale_x_continuous(
    name = "Attachment Avoidance" 
  ) + 
  scale_y_continuous(
    name = "Influence of Satisfaction"
  )

plot_satisfaction

ggsave("plot_satisfaction.pdf")

```


Caveat prior to explaining graphs - these graphs are estimates. I could have fit a spline or something more complicated to better understand the model, and then graphed the model in a way that told a more clear story, but that is something that will take more time and practice. My explanations for the graphs will be based on the raw data and estimates. 

From the above graph, you can see that the negative relationship between attachment avoidance and perceived influence of satisfaction seems to only hold true for those low in attachment anxiety. It also seems that there as an opposite or increasing tend (however slight) when individuals with high attachment avoidance are also high in attachment anxiety. 

## Attachment predicting mutuality

### model 

```{r fit mod mutuality}
m_mutuality <- lm(
  mutuality_1 ~ AnxiousTotalMean + AvoidTotalMean + AnxiousTotalMean*AvoidTotalMean, 
  data = dat_layperson
)
``` 

### parameters 

```{r parameters m_mutuality}

params_mutuality<- parameters( 
  m_mutuality, 
  dispersion = TRUE, 
  component = "all", 
  test = NULL, 
  diagnostic = NULL, 
  prior = NULL 
) 
 
print_md(params_mutuality) 

```

### Findings

According to the parameters, anxious attachment style is a negative predictor of perception of mutuality's influence on relationship health. Every 1 unit increase in attachment anxiety corresponds to a -4.20 decrease in perceived healthiness of satisfaction. There was a moderately sized compatibility interval that does not include 0 [-7.38, -1.03], meaning there is evidence that this relationship would be seen in the population. Avoidant attachment style is also a negative predictor of perception of mutuality on relationship healthy. Every 1 unit increase in attachment avoidance corresponds to a -11.13 decrease in perceived healthiness of satisfaction. The compatibility interval [-15.89, -6.38] does not include 0, and we would not be surprised to see this relationship in the population. Interestingly, there was a significant interaction, CI [0.96, 3.42], that trended positively (2.19). Graphing this data will provide more information on the relationship. 

### plot

```{r mod mut plot}
plot_mutuality <- dat_layperson1 |>
  filter(!is.na(AnxiousTotalMeanInt)) |>
  ggplot() +
  aes(x = AvoidTotalMean, 
      y = mutuality_1, 
      group = AnxiousTotalMeanInt, 
      color = AnxiousTotalMeanInt) +
  geom_point(shape=21,
             color="darkgrey",
             fill= "darkgrey", 
             size=2, 
             alpha = .5) +
  geom_smooth(method=lm) +
  scale_x_continuous(
    name = "Attachment Avoidance" 
  ) + 
  scale_y_continuous(
    name = "Influence of Mutuality"
  )

plot_mutuality

ggsave("plot_mutuality.pdf")
```

From the above graph, you can see that the negative relationship between attachment avoidance and influence of mutuality also seems to only hold for those who are low in anxiety. This, coupled with the graph for satisfaction, seems to insinuate that anxiety buffers the effect of avoidance. 


## Attachment predicting caring

### model 

```{r fit mod caring}
m_caring <- lm(
  caring_1 ~ AnxiousTotalMean + AvoidTotalMean + AnxiousTotalMean*AvoidTotalMean, 
  data = dat_layperson
)
``` 

### parameters

```{r parameters m_caring}

params_caring <- parameters( 
  m_caring, 
  dispersion = TRUE, 
  component = "all", 
  test = NULL, 
  diagnostic = NULL, 
  prior = NULL 
) 
 
print_md(params_caring) 

```

### Findings

According to the parameters, anxious attachment style is a negative predictor of perception of caring's influence on relationship health. Every 1 unit increase in attachment anxiety corresponds to a -1.97 decrease in perceived healthiness of satisfaction. There was a moderately sized compatibility interval that does not include 0 [-3.81, -0.13], meaning there is evidence that this relationship would be seen in the population. Avoidant attachment style is also a negative predictor of perception of caring on relationship healthy. Every 1 unit increase in attachment avoidance corresponds to a -5.63 decrease in perceived healthiness of satisfaction. The compatibility interval [-8.41, -2.86] does not include 0, and we would not be surprised to see this relationship in the population. Interestingly, there was a significant interaction, CI [0.17, 1.61], that trended positively (0.37). Graphing this data will provide more information on the relationship. 

### plot

```{r mod caring plot}
plot_caring <- dat_layperson1 |>
  filter(!is.na(AnxiousTotalMeanInt)) |>
  ggplot() +
  aes(x = AvoidTotalMean, 
      y = mutuality_1, 
      group = AnxiousTotalMeanInt, 
      color = AnxiousTotalMeanInt) +
  geom_point(shape=21,
             color="darkgrey",
             fill= "darkgrey", 
             size=2, 
             alpha = .5) +
  geom_smooth(method=lm) +
  scale_x_continuous(
    name = "Attachment Avoidance" 
  ) + 
  scale_y_continuous(
    name = "Influence of Caring"
  )

plot_caring

ggsave("plot_caring.pdf")
```

From the above graph, there is a similar pattern as the other two already, where the strong negative relationship is only holding for low anxiety participants. 

## Attachment predicting core similarities 

### model 

```{r fit mod core sim}
m_coresim <- lm(
  coresim_1 ~ AnxiousTotalMean + AvoidTotalMean + AnxiousTotalMean*AvoidTotalMean, 
  data = dat_layperson
)
``` 

### parameters

```{r parameters m_coresim}

params_coresim <- parameters( 
  m_coresim, 
  dispersion = TRUE, 
  component = "all", 
  test = NULL, 
  diagnostic = NULL, 
  prior = NULL 
) 
 
print_md(params_coresim) 

```

### Findings

According to the parameters, anxious attachment style is a negative predictor of perception of core similarities' influence on relationship health. Every 1 unit increase in attachment anxiety corresponds to a -2.30 decrease in perceived healthiness of core similarities. There was a moderately sized compatibility interval that does not include 0 [-4.55, -0.05], meaning there is evidence that this relationship would be seen in the population. Avoidant attachment style is also a negative predictor of perception of caring on relationship healthy. Every 1 unit increase in attachment avoidance corresponds to a -6.37 decrease in perceived healthiness of core similarities. The compatibility interval [-9.77, -2.96] does not include 0, and we would not be surprised to see this relationship in the population. Interestingly, there was a significant interaction, CI [0.36, 2.12], that trended positively (0.45). Graphing this data will provide more information on the relationship. 

### plot

```{r mod coresim plot}
plot_coresim <- dat_layperson1 |>
  filter(!is.na(AnxiousTotalMeanInt)) |>
  ggplot() +
  aes(x = AvoidTotalMean, 
      y = coresim_1, 
      group = AnxiousTotalMeanInt, 
      color = AnxiousTotalMeanInt) +
  geom_point(shape=21,
             color="darkgrey",
             fill= "darkgrey", 
             size=2, 
             alpha = .5) +
  geom_smooth(method=lm) +
  scale_x_continuous(
    name = "Attachment Avoidance" 
  ) + 
  scale_y_continuous(
    name = "Influence of Core Similarities"
  )

plot_coresim

ggsave("plot_coresim.pdf")
```


Lastly, there is a similar pattern here (but seems less strong than the other graphs)

This was exploratory data and we do not yet have hypotheses about why these interactions were seen. Also, there is a clear ceiling effect, where most participants rated qualities as making relationships healthy. 